Journal and article coverage

### 2024-08-20 analysis/results.Rmd
Style upset plots

#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  billing = "subugoe-collaborative",
  project = "subugoe-collaborative"
)
```

@@

# Backup
write_csv(jn_upset_articles, here::here("data/jn_upset_articles.csv"))
```

```{r journal_article_overlap_upset}
size <- get_size_mode("exclusive_intersection")

upset_overview <- ComplexUpset::upset(
  jn_upset_articles,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  width_ratio = 0.4,
  set_sizes = FALSE,
  stripes = "transparent",
 # wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = TRUE,
        text = list(size = 9 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        )
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ","))
    )
  ),
  annotations = list("Journal size\n(log scale)" = (
    ggplot(mapping = aes(x = intersection, y = n)) +
      geom_boxplot(outliers = FALSE) +
      scale_y_log10(labels =  scales::number_format(big.mark = ","))),
    "Articles" = (ggplot(mapping = aes(y = n_cr)) +
                        geom_bar(stat = "identity", color = "grey80", aes(fill = "in Crossref only")) +
                        geom_bar(aes(y = doi_in_wos_and_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
                        geom_bar(aes(y = doi_in_wos, fill = "Shared"), stat = "identity", color = "#0093c7") +
                        geom_bar(aes(y = doi_in_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
                        scale_fill_manual(values = c(`in Crossref only` = "grey80", `Shared` = "#0093c7"), name = "") +
                        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
                        theme(
    legend.position = "top",
    legend.justification = "right"
  ) +
    guides(fill = "none")
    )),
   matrix = (
        intersection_matrix(
            geom = geom_point(
                shape = "square",
                size = 2
            )))
        ) 
upset_overview
```
The plot is limited to journals, which are indexed in Crossref. However, there
are `r jn_upset_articles |> filter(Crossref == FALSE) |> distinct(issn_l) |> nrow()` journals not incldued in Crossref, but in WoS or Scopus:

@@

```{r upset_publisher_plot}
size = get_size_mode('exclusive_intersection')

upset_publisher_plot <- ComplexUpset::upset(
  jn_upset_publisher,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  width_ratio = 0.4,
  stripes = "transparent",
  name = "Data Source Coverage",
  base_annotations = NULL,
  annotations = list(
    "Journal size\n(log-scale)" = (
      ggplot(mapping = aes(x = intersection, y = n, fill = fct_rev(publisher))) +
        geom_boxplot(outliers = FALSE) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none") +
        scale_y_log10(labels = scales::number_format(big.mark = ",")) 
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n, fill = publisher, color = publisher)) +
        geom_bar(stat = "identity") +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3a0"
          ), guide = "none") +
        scale_color_manual(
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none"
        ) + scale_y_continuous(labels = scales::number_format(big.mark = ",")) 
    ),
    "Publisher" = (
      ggplot(mapping = aes(fill = publisher)) +
        geom_bar(stat = "count", position = "fill") + 
        scale_y_continuous("Journals", labels = scales::percent_format()) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none")
    )
  ),
  set_sizes = (
    upset_set_size(
      geom = geom_bar(aes(fill = publisher), width = 0.8), 
      position = "right"
    ) +
    scale_fill_manual(
      "",
      values = c(
        "Elsevier" = "#e9711c",
        "Springer Nature" = "#486a7e",
        "Wiley" = "#068853",
        "Other" = "#b3b3b3"
      )
    ) +
    scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
    guides(fill = guide_legend(reverse = TRUE)) +
    ylab("Journals")
  ),
  guides = "over",
  matrix = (
    intersection_matrix(
      geom = geom_point(
        shape = "square",
        size = 2
      )
    )
  )
)

upset_publisher_plot 
```

>>>>

#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  billing = "subugoe-collaborative",
  project = "subugoe-collaborative"
)

# Set themes for plots
my_base_size = 12

my_upset_theme <-
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        plot.title.position="plot",
        plot.title = element_text(hjust = 0, vjust = 1, face = "bold"))

```

@@

# Backup
jn_upset_articles
write_csv(jn_upset_articles, here::here("data/jn_upset_articles.csv"))
```

```{r journal_article_overlap_upset, fig.asp=0.75}
# Helper function to make text labels the same size as axis labels as seen in
# <https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/>
# Todo: Learn how to invoke it
ggtext_size <- function(base_size, ratio = 0.8) {
  ratio * base_size / ggplot2::.pt
}

bar_width <- 0.7
size <- get_size_mode("exclusive_intersection")

upset_overview <- ComplexUpset::upset(
  jn_upset_articles,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  set_sizes = FALSE,
  stripes = "transparent",
  # wrap = TRUE,
  name = "Data Source Coverage",
  base_annotations = list(
    "Journals" = (
      ComplexUpset::intersection_size(
        counts = TRUE,
        text = list(size = 8 / .pt),
        text_mapping = aes(
          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
          y = ifelse(!!size > 5000, !!size - 4000, !!size)
        ), width = bar_width
      ) +
        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
        ggtitle("A")
    )
  ),
  annotations = list(
    "Journal size\n(log scale)" = (
      ggplot(mapping = aes(x = intersection, y = n)) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n_cr)) +
        geom_bar(
          stat = "identity",
          color = "grey80",
          width = bar_width,
          aes(fill = "in Crossref only")
        ) +
        geom_bar(
          aes(y = doi_in_wos_and_scopus, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0093c7"
        ) +
        geom_bar(
          aes(y = doi_in_wos, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0093c7"
        ) +
        geom_bar(
          aes(y = doi_in_scopus, fill = "Shared"),
          stat = "identity",
          width = bar_width,
          color = "#0093c7"
        ) +
        scale_fill_manual(
          values = c(
            `in Crossref only` = "grey80",
            `Shared` = "#0093c7"
          ),
          name = "DOI"
        ) +
        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
        ggtitle("B")
    )
  ),
  matrix = (intersection_matrix(geom = geom_point(
    shape = "square", size = 2
  )))
)
upset_overview & my_upset_theme
```

The plot is limited to journals, which are indexed in Crossref. However, there
are `r jn_upset_articles |> filter(Crossref == FALSE) |> distinct(issn_l) |> nrow()` journals not incldued in Crossref, but in WoS or Scopus:

@@

```{r upset_publisher_plot, fig.asp=0.75}
size = get_size_mode('exclusive_intersection')

upset_publisher_plot <- ComplexUpset::upset(
  jn_upset_publisher,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  width_ratio = 0.4,
  stripes = "transparent",
  name = "Data Source Coverage",
  base_annotations = NULL,
  annotations = list(
    "Journal size\n(log-scale)" = (
      ggplot(mapping = aes(x = intersection, y = n, fill = fct_rev(publisher))) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none") +
        scale_y_log10(labels = scales::number_format(big.mark = ",")) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = n, fill = publisher, color = publisher)) +
        geom_bar(stat = "identity", width = bar_width) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3a0"
          ), guide = "none") +
        scale_color_manual(
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none"
        ) + scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
        ggtitle("B") 
    ),
    "Publisher" = (
      ggplot(mapping = aes(fill = publisher)) +
        geom_bar(stat = "count", position = "fill", width = bar_width) + 
        scale_y_continuous("Journals", labels = scales::percent_format()) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none") +
        ggtitle("A")
    )
  ),
  set_sizes = (
    upset_set_size(
      geom = geom_bar(aes(fill = publisher), width = 0.8), 
      position = "right"
    ) +
    scale_fill_manual(
      "",
      values = c(
        "Elsevier" = "#e9711c",
        "Springer Nature" = "#486a7e",
        "Wiley" = "#068853",
        "Other" = "#b3b3b3"
      )
    ) +
    scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
    guides(fill = guide_legend(reverse = TRUE)) +
    ylab("Journals")
  ),
  guides = "over",
  matrix = (
    intersection_matrix(
      geom = geom_point(
        shape = "square",
        size = 2
      )
    )
  )
)

upset_publisher_plot & my_upset_theme 
```
